Sentinel lead pipe racks quantify orthophosphate's dose-response in drinking water

### 2023-05-26 README.Rmd
first journal submission

This repository contains the code and data necessary to reproduce the main results of the preprint ["Sentinel pipe racks quantify orthophosphate's effect on lead release into drinking water"](https://doi.org/10.26434/chemrxiv-2023-h1459).

>>>>

This repository contains the code and data necessary to reproduce the main results of the preprint ["Sentinel lead pipe racks quantify orthophosphate’s dose-response in drinking water"](https://doi.org/10.26434/chemrxiv-2023-h1459).

### 2023-05-26 Rmarkdown/paper-si.Rmd
first journal submission

---
title: "*Supplementary information for:* Sentinel pipe racks quantify orthophosphate's effect on lead release into drinking water"
author: "Benjamin F. Trueman^§,^\\*, Wendy H. Krkošek^†^, and Graham A. Gagnon^§^"
date: "^§^Centre for Water Resources Studies, Department of Civil & Resource Engineering, Dalhousie University, 1360 Barrington St., Halifax, Nova Scotia, Canada B3H 4R2  \n&nbsp;  \n^†^Halifax Water, 450 Cowie Hill Rd., Halifax, Nova Scotia, Canada,  B3P 2V3  \n&nbsp;   \n&nbsp;*Corresponding author  \n&nbsp;E-mail: benjamin.trueman@dal.ca    \n&nbsp;Tel:902.494.6070  \n&nbsp;Fax: 902.494.3105"
output:
  word_document:
    reference_docx: StylesTemplate.docx
bibliography: references.bib
csl: environmental-science-and-technology.csl
---

>>>>

---
title: "*Supplementary information for:* Sentinel lead pipe racks quantify orthophosphate's dose-response in drinking water"
author: "Benjamin F. Trueman^§,^\\*, Wendy H. Krkošek^†^, and Graham A. Gagnon^§^"
date: "^§^Centre for Water Resources Studies, Department of Civil & Resource Engineering, Dalhousie University, 1360 Barrington St., Halifax, Nova Scotia, Canada B3H 4R2  \n&nbsp;  \n^†^Halifax Water, 450 Cowie Hill Rd., Halifax, Nova Scotia, Canada,  B3P 2V3  \n&nbsp;   \n&nbsp;*Corresponding author  \n&nbsp;E-mail: benjamin.trueman@dal.ca    \n&nbsp;Tel:902.494.6070  \n&nbsp;Fax: 902.494.3105"
output:
  word_document:
    reference_docx: StylesTemplate.docx
bibliography: references.bib
csl: environmental-science-and-technology.csl
---

### 2023-05-26 Rmarkdown/paper.Rmd
first journal submission

We used R [@R], and a collection of contributed packages [@rmarkdown; @tidyverse; @readxl; @janitor; @assertr; @brms; @ggdist; @glue; @posterior; @testthat; @bayesplot; @patchwork; @ggtext; @devtools; @mgcv], to analyze and visualize the data. The code and data necessary to reproduce the main results of the paper are available online [@cct_monitoring].

@@

$$
(2)~~
\begin{align}
\text{likelihood:} \\
log(y_i) | censored = 0 \sim T(\mu_i, \sigma) \\
log(y_i) | censored = 1 \sim T\text{-}CDF(\mu_i, \sigma) \\
\\
\text{model for }\mu_i\text{:} \\
\mu_i = \alpha_{site_j} + \gamma_{sample_k} + \beta R \\
\\
\text{priors:} \\
\sigma \sim Half \text{-} N(0,1) \\
\nu \sim Gamma(2, 0.1) \\
\beta \sim N(-0.8,0.3) \\
\\
\alpha_{site_j} \sim N(\bar{\alpha}, \sigma_{\alpha}),~\text{for}~j~\text{in}~1..9 \\
\bar{\alpha} \sim N(0,1) \\
\sigma_{\alpha} \sim Half\text{-}Cauchy(0,1) \\
\\
\gamma_{sample_k} \sim N(0, \sigma_{\gamma}),~\text{for}~k~\text{in}~1..45 \\
\sigma_{\gamma} \sim Half\text{-}Cauchy(0,1) \\
\end{align}
$$

@@

The priors on $\bar{\alpha}$, $\bar{\gamma}$, $\sigma_{\alpha}$ , and $\sigma_{\gamma}$ are weakly informative, meaning that they discourage unrealistic parameter estimates [@gelman_bayes_2009]. The prior on $\beta$---the difference between lead concentration at the two orthophosphate doses---was determined using posterior predictions from the generalized additive model of pipe loop data, as described in the results section.

>>>>

We used R [@R], and a collection of contributed packages [@rmarkdown; @tidyverse; @readxl; @janitor; @assertr; @brms; @ggdist; @glue; @posterior; @testthat; @bayesplot; @patchwork; @ggtext; @devtools; @mgcv], to analyze and visualize the data. Materials (R code and data) necessary to reproduce the main results of the paper are available online [@cct_monitoring].

@@

$$
(2)~~
\begin{align}
\text{likelihood:} \\
log(y_i) | censored = 0 \sim T(\mu_i, \sigma, \nu) \\
log(y_i) | censored = 1 \sim T\text{-}CDF(\mu_i, \sigma, \nu) \\
\\
\text{model for }\mu_i\text{:} \\
\mu_i = \alpha_{site_j} + \gamma_{sample_k} + \beta R \\
\\
\text{priors:} \\
\sigma \sim Half \text{-} N(0,1) \\
\nu \sim Gamma(2, 0.1) \\
\beta \sim N(-0.8,0.3) \\
\\
\alpha_{site_j} \sim N(\bar{\alpha}, \sigma_{\alpha}),~\text{for}~j~\text{in}~1..9 \\
\bar{\alpha} \sim N(0,1) \\
\sigma_{\alpha} \sim Half\text{-}Cauchy(0,1) \\
\\
\gamma_{sample_k} \sim N(0, \sigma_{\gamma}),~\text{for}~k~\text{in}~1..45 \\
\sigma_{\gamma} \sim Half\text{-}Cauchy(0,1) \\
\end{align}
$$

@@

The priors on $\bar{\alpha}$, $\sigma_{\alpha}$ , and $\sigma_{\gamma}$ are weakly informative, meaning that they discourage unrealistic parameter estimates [@gelman_bayes_2009]. The prior on $\beta$---the difference between lead concentration at the two orthophosphate doses---was determined using posterior predictions from the generalized additive model of pipe loop data, as described in the Results and discussion.
